\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[a4paper, total={6in, 8in}]{geometry}
\usepackage{indentfirst}

\usepackage{todonotes}

\title{CS5199: Individual Masters Project\\DOER}
\author{Thomas E. Hansen (150015673)\\Supervisor: Dr. John Thomson}
\date{Feb. 2020}

\begin{document}
    
    \maketitle
    
    \section{Description}
    \begin{itemize}
        \item something about scheduling
        \item something about ARM? big.LITTLE? Mobile processors?
        \item something about power consumption?
        \item something about dynamic frequency scaling?
    \end{itemize}
    
    \section{Objectives}
        \subsection{Primary}
        \begin{itemize}
            \item Become familiar with the gem5 simulator
            \item Use gem5 to explore the gains that dynamic voltage regulation
                  may bring to certain tasks
            \item Examine how this may be done automatically in a scheduler
        \end{itemize}
        
        \subsection{Secondary}
        \begin{itemize}
            \item Implement the scheduling heuristics in the Linux scheduler
            \item Analyse the effects of the new scheduler using the gem5
                  simulator
        \end{itemize}
        
        \subsection{Ternary}
        \begin{itemize}
            \item Analyse the effects of the new scheduler using an ARM
                  development board, i.e. on real hardware
        \end{itemize}
        
    \section{Ethics}
    This project has no ethical concerns. The project will not deal with
    personal data. Any data used will be entirely synthetic or public domain.
    
    \section{Resources}
    This project does not require any exceptional resources.
\end{document}
