\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[a4paper, total={6in, 8in}]{geometry}
\usepackage{indentfirst}

\usepackage{hyperref}
\usepackage{url}

\hypersetup{colorlinks=True
            , urlcolor=blue
           }

\title{CS5199: Individual Masters Project\\DOER}
\author{\textbf{Student:} Thomas E. Hansen (150015673)\\\textbf{Supervisor:}
    Dr. John Thomson}
\date{Feb. 2020}

\begin{document}
    
    \maketitle
    
    \section{Description}
    This project aims to investigate Dynamic Voltage Loop Scheduling (DVLS) for
    ARM ``big.LITTLE'' processors. This will be done by using the
    \href{http://www.gem5.org/Main_Page}{gem5 simulator} to measure energy
    consumption and performance impacts. Additionally, an ARM development board
    will be used to test the software developed on real hardware. The project
    will involve developing new methods for adjusting the voltage of CPUs in a
    runtime, e.g. the Linux scheduler or a governor.
    
    \section{Objectives}
        \subsection{Primary}
        \begin{itemize}
            \item Become familiar with the gem5 simulator.
            \item Use gem5 to explore the impact that DVLS may have on certain
                  tasks.
            \item Examine how DVLS may be done automatically in a runtime.
        \end{itemize}
        
        \subsection{Secondary}
        \begin{itemize}
            \item Implement the DVLS heuristics in the Linux scheduler or
                  implement a governor.
            \item Analyse the effects of the new runtime using the gem5
                  simulator.
        \end{itemize}
        
        \subsection{Ternary}
        \begin{itemize}
            \item Analyse the effects of the new runtime using an ARM
                  development board.
        \end{itemize}
        
    \section{Ethics}
    This project has no ethical concerns. The project will not deal with
    personal data. Any data used will be entirely synthetic or public domain.
    
    \section{Resources}
    This project will require an ARM development board costing about £50-£100.
\end{document}
